Long name of the paper

### 2018-10-29 anzjs2018/sections/04_implementation.tex
more writing and pictures

The parameters varied were the number of particles,
GPS error $\epsilon$, and system noise $\sigma^2$.
The primary goal is to reduce the rate of degeneration and ``losing'' the bus 
(i.e., no particles are near the observed location),
and reducing the variance of parameter estimates; here we use vehicle speed as the example.

The one parameter we can determine from the data is GPS error, $\epsilon^2$,
by looking at the distance between observations and the shape.
Figure~\ref{fig:gps_dist} shows the distribution of distance between the vehicle
observation and the \emph{nearest point on the path}.
Due to the way the observations are reported, 
many observations will be ``exactly'' on the route at a bus stop,
and we need to ignore these.

To check the model results, 
we need to realise that we are not trying to accurately predict
future locations of the vehicle,
but instead are trying to use a sequence of GPS positions to infer travel
times along intermediate roads.
Therefore, several tools used to assess the validity of a particle filter
are unnecessary.
Instead, we will be comparing the rate of degeneration 
(i.e., the particle sample loses the actual bus and needs to be reinitialized)
and the mean distance between the state estimate and the observation; 
we will use the distance between the observation and the route (cross track distance) 
to determine how well the posterior distribution fits the data---%
if the ratio between the CTD and RMSE is large,
then the fit is ``bad''.
[We also explore the uncertainty of travel time estimates,
as well as the variability between consecutive buses]\footnote{assuming we can get the results in time}.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.7\textwidth]{figures/04_model_results_dist.pdf}
    \caption{The timings for various parts of the application, and overall. %
        The trend is approximately linear with the number of particles.}
    \label{fig:dist_to_route}
\end{figure}

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.7\textwidth]{figures/04_model_results_neff.pdf}
    \caption{The timings for various parts of the application, and overall. %
        The trend is approximately linear with the number of particles.}
    \label{fig:neff}
\end{figure}

From the Figure~X, 
we see a trade-off between some of the parameters.
Larger values of GPS error decrease the degeneration rate
as well as the rate of resampling (which increases speed);
however, the cost of this is that the mean distance between the posterior distribution
and the observed location increases, giving a less precise estimate of the vehicle's state.
Conversely, the system noise does not have the same effect,
and only slightly effects the values.

% How we implement it, choice of software (Rcpp = R + C++).
% R: dealing with data structures is easier, maintainability, interfacing
% C++: speed

% Overall structure:
% - load
% - fetch positions
% - initialize or mutate+update
% - update network
% - make ETA predictions (vehicle state + network state)

% Some of the key things:
% - minimise copy, parallelisation using OMP
% - moving as much computation ``outside'' of the main loop as possible
%   (e.g., ``pre''-predict vehicle/network states so only update required)
%   so ETAs are generated ASAP after retrieving data
% - keeping the GTFS database up-to-date by fetching new data each morning
% - distribution - a cloud database vs a single protobuf file with everything 
%   (maintenance/reliability/speed/size)

>>>>

The model was evaluated using several statistics,
each calculated using a range of parameter values.
The first is effective sample size,
defined in (\ref{eq:neff}),
which is used to determined if resampling is required.
Another important statistic is the \emph{degeneration rate},
or the percent of iterations where the particle filter
is unable to find any plausible states given the observation,
and the vehicle is reinitialised.
Note, however, that the lower limit of this value 
is attributed to \emph{invalid trips},
however there is no simple way of filtering these trips.
The last statistic we compute is the variance of road segment travel times
over time.
For this example, we varied the number of particles $N$,
GPS error $\epsilon$, and system noise $\sigma^2$.

The one parameter we can determine from the data is GPS error, $\epsilon^2$,
by looking at the distance between observations and the shape.
Figure~\ref{fig:dist_to_route} shows the distribution of distance between the vehicle
observation and the \emph{nearest point on the path}.
Due to the way the observations are reported, 
many observations will be ``exactly'' on the route at a bus stop,
which explains the spike at just above zero~meters.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.7\textwidth]{figures/04_model_results_dist.pdf}
    \caption{The distance between the observation and the nearest point on the shape.}
    \label{fig:dist_to_route}
\end{figure}

For the effective sample size, 
increasing GPS error and system error both tend to
increase $N_\text{eff}$, as shown in Figure~\ref{fig:degen_rate}.
Increasing the GPS error distributes the weights more evenly,
while increasing system noise allows more particles to reach
the necessary speed,
so these are as expected.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/04_model_results_degen.pdf}
    \caption{The effective sample size and degeneration rate for a range of parameter 
        values. The columns are of increasing number of particles,
        and the rows are for increasing (downwards) GPS error.}
    \label{fig:degen_rate}
\end{figure}

The degeneration rate, 
shown on the $y$-axis of Figure~\ref{fig:degen_rate},
decreases as GPS error increases,
as anomolous observations where the bus is far from the route
are more likely.
It also decreases with the number of particles,
while system noise has no apparent effect.

Lastly is the variability of travel times and their uncertainties.
These are shown in Figure~\ref{fig:travel_times}.
System noise again shows no effect on the variability of travel time estimates,
however increasing GPS error tends to increase travel time variability.
This means that there is a trade-off between the models performance
and how well it estimates the main parameter of interest:
larger values of $\epsilon$ result in less degeneration and larger $N_\text{eff}$,
at the cost of more variable travel time estimates.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/04_model_results_times.pdf}
    \caption{Travel times.}
    \label{fig:travel_times}
\end{figure}

### 2018-10-29 anzjs2018/sections/06_discuss.tex
more writing and pictures

It all seems to work (hopefully I don't need to change this D:)

Next steps include
- generating useful priors for ``fall back'' (i.e., no data, future) prediction
- develop and implement a more complex network update model (correlations, etc)
- prediction estimates (point vs interval)

>>>>

The main goal of this work was to explore the faesibility
of a particle filter as a \rt estimation method for transit vehicle states,
and the use of estimated states to determine traffic conditions
throughout the network.
This allows vehicles to inform others---irrespective of route---of 
how long it will take to reach future stops.

Not only is the particle filter a viable choice in terms of speed,
it has shown it can handle most situations without losing the vehicle,
with the exception of when the vehicle is not traveling along
the specified route.
Hopefully with future development,
the stopping behaviour can be more accurately modeled,
perhaps incorporating trip updates into the likelihood function.

The network model is still in the early stages of development,
but given that the implementation is running quickly---twice as fast as our target---%
there is room for developing a more complex model.
Future work will explore the use of historical data to develop informative priors
for travel times,
which will provide a useful fallback in the case of no data,
or for making more accurate long-term predictions.
Once such a model has been developed, we plan to explore prediction estimates,
particuarly comparing point and intervals as a means of communicating 
uncertainty to commuters.
